Clicks and Cliques

### 2018-12-28 papersoul.Rnw
small edits

\begin{abstract} 
In the  paper we analyze 26 communities across the United States with the objective to understand what attaches people to their community and how this attachment differs among communities. How different are attached people from unattached? What attaches people to their community? How different are the communities? What are key drivers behind emotional attachment? To address these questions, graphical, supervised and unsupervised learning tools were used and information from the Census Bureau and the Knight Foundation were combined. 
Using the same pre-processed variables as \cite{full2010} most likely will drive the results towards the same conclusions than the Knight foundation, so this paper does not use those variables.

@@

complete <- 100*sum(miss.c2 == 0)/length(miss.c2)
dim.case <- dim(soul)[1]*sum(miss.c2 == 0)/length(miss.c2)
dim.var <- sum(miss.v < .25)
@

In order to obtain a data set for the statistical analysis part of the information is not considered. First, only variables within the original and demographic group are used.
Secondly, the set of variables with less than 25\% of missing cases are selected. Among the selected variables, only complete cases are used. The final data set created with this procedure produce 30000 complete observations and 65 variables.

\subsection{Statistical methods}

Supervised and unsupervised learning tools were used in this paper to answer the main questions presented in Section \ref{intro}. To present the results of statistical learning methods, different visualizations were used to perform an exploratory analysis. In this section the main tools for graphical and statistical analysis are described.

\subsubsection{Sampling weights}
Statistical analysis of data from survey need to make a correct use of sampling weights. Depending of the source of information the type of statistical goal, different sampling weights are used in several statistical methods or summaries.

@@

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that should be used depending if data from many communities are used jointly (PROJWT) or computations are done for each community separatedly (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities toghether so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separatedly within each community  WEIGHT variable is used as sampling weight. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be consider an estimate of the comunity total.

@@

Cluster analysis is an unsupervised technique and two ways to find clusters are used across this paper. Hierarchical agglomerative cluster is used to find clusters at the community level. This clustering algorithm starts with each point being its own cluster and is recursively merged. Then, the ``merging history'' forms a hierarchical tree diagram called a dendrogram, which is a tool used to define the number of clusters (see Sec. 12.3.1 of \cite{izenman2008modern} for more details). Additionally, K-means clustering was used to find clusters at the individual level. The K-means algorithm, proposed by \cite{macqueen1967some}, starts with all observations assigned to $K$ initial clusters and cluster means are computed. Then in an iterative fashion observations are reassigned to minimize the within sums of squares of each cluster until a criterion of convergence has been met (see Sec. 12.4.1 of \cite{izenman2008modern} for more details).

@@

\subsubsection{Random forest}
Random forest is a supervised learning technique proposed by \cite{breiman2001random}, based on combining results from many tree predictors computed on bootstrap samples of the data. A tree predictor is a classifier which recursively partition the feature space looking for regions where the response variable is homogeneous (pure nodes). In order to build a forest, bootstrap samples from the data are selected and for each sample a tree classifier is grown. However, the constructed tree in each bootstrap sample only uses a random selection of variables to find the next feature partition. Then two sources of randomness are used, random selection of cases in the bootstrap sample and random selection of variables. Then the forest classifier is an ensemble of randomized trees, the forest prediction is determined with majority vote across trees (see Section 14.4 of \cite{izenman2008modern} for more details).

@@

The most common use for the random forest classifier is for prediction, however the main objective is to identify the key variables to explain the attachment for each community instead of predict the attachment level.

@@

There are two basic measures of variable importance in the \verb randomForest  function within R, only the \emph{mean decrease in Gini} ($DG$) was used. Each time a variable is used within a random forest classifier is associated with a decrease in the impurity of the data nodes, which is a measure with the Gini impurity index. Important variables will achieve larger reduction of impurity, then $DG$ can be used as a variable importance measurement.

@@

Within each community explanatory variables were ranked according to $DG$. Rank 1 represents the variable associated with the largest reduction in the data impurity. Figure \ref{imp1} shows variable importance, and color intensity indicates how many times a variable appears in each of the first five ranks. At a glance it is clear that 5.year.ahead, 5.year.ago and Meet.People, appears as the most important variables in most of the communities. The variable 5.year.ahead indicates the importance of forward looking, i.e. what do you expect from your community in 5 years, and this is the most important driver of attachment.

The variable 5.years.ago represents the perception of the community improvement, when comparing the community today with 5 years ago. Finally, the Meet.People variable indicates if the community is a good place to meet people and make friends.
\begin{figure}[hbpt]
<< import.plot3, dependson=c('data'), echo=FALSE >>=
aux <- soul.red2
aux$QSB <- with(aux, reorder(QSB, -as.numeric(QSB) ))

>>>>

\begin{abstract} 
In this  paper we analyze 26 communities across the United States with the objective to understand what attaches people to their community and how this attachment differs among communities. How different are attached people from unattached? What attaches people to their community? How different are the communities? What are key drivers behind emotional attachment? To address these questions, graphical, supervised and unsupervised learning tools were used and information from the Census Bureau and the Knight Foundation were combined. 
Using the same pre-processed variables as \cite{full2010} most likely will drive the results towards the same conclusions than the Knight foundation, so this paper does not use those variables.

@@

complete <- 100*sum(miss.c2 == 0)/length(miss.c2)
dim.case <- dim(soul)[1]*sum(miss.c2 == 0)/length(miss.c2)
dim.var <- sum(miss.v < .25)
@
In order to obtain a data set for the statistical analysis part of the information is not considered. First, only variables within the original and demographic group are used.
Secondly, the set of variables with less than 25\% of missing cases are selected. Among the selected variables, only complete cases are used. The final data set created with this procedure produce 30000 complete observations and 65 variables.

\subsection{Statistical methods}
Supervised and unsupervised learning tools were used in this paper to answer the main questions presented in Section \ref{intro}. To present the results of statistical learning methods, different visualizations were used to perform an exploratory analysis. In this section the main tools for graphical and statistical analysis are described.

\subsubsection{Sampling weights}
Statistical analysis of data from survey need to make a correct use of sampling weights. Depending of the source of information and the statistical goal, different sampling weights are used in several statistical methods or summaries.

@@

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that might be used considering if data from many communities are combined (PROJWT) or computations are done for each community separatedly (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities toghether so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separatedly within each community  and therefore sampling weights correspond to WEIGHT variable. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be consider an estimate of the comunity total.

@@

Cluster analysis is an unsupervised technique, two ways to find clusters are used across this paper. Hierarchical agglomerative cluster is used to find clusters at the community level. This clustering algorithm starts with each point being its own cluster and is recursively merged. Then, the ``merging history'' forms a hierarchical tree diagram called a dendrogram, which is a tool used to define the number of clusters (see Sec. 12.3.1 of \cite{izenman2008modern} for more details). Additionally, K-means clustering was used to find clusters at the individual level. The K-means algorithm, proposed by \cite{macqueen1967some}, starts with all observations assigned to $K$ initial clusters and cluster means are computed. Then in an iterative fashion observations are reassigned to minimize the within sums of squares of each cluster until a criterion of convergence has been met (see Sec. 12.4.1 of \cite{izenman2008modern} for more details).

@@

\subsubsection{Random forest}
Random forest is a supervised learning technique proposed by \cite{breiman2001random}, based on combining results from many tree predictors computed on bootstrap samples of the data. A tree predictor is a classifier which recursively divides the feature space looking for regions where the response variable is homogeneous (pure nodes). In order to build a forest, bootstrap samples from the data are selected and for each sample a tree classifier is grown. However, the constructed tree in each bootstrap sample only uses a random selection of variables to find the next feature partition. Then two sources of randomness are used, random selection of cases in the bootstrap sample and random selection of variables. Finally, the forest classifier is an ensemble of randomized trees, the forest prediction is determined with majority vote across trees (see Section 14.4 of \cite{izenman2008modern} for more details).

@@

The most common use for the random forest classifier is for prediction, however in this paper the main objective is to identify the key variables to explain the attachment for each community instead of predict the attachment level.

@@

There are two basic measures of variable importance in the \verb randomForest function within R, only the \emph{mean decrease in Gini} ($DG$) was used. Each time a variable is used within a random forest classifier is associated with a decrease in the impurity of the data nodes, which is a measure with the Gini impurity index. Important variables will achieve larger reduction of impurity, then $DG$ can be used as a variable importance measurement. Within each community explanatory variables were ranked according to $DG$. Rank 1 represents the variable associated with the largest reduction in the data impurity.

@@

Figure \ref{imp1} shows variable importance, and color intensity indicates how many times a variable appears in each of the first five ranks. At a glance it is clear that 5.year.ahead, 5.year.ago and Meet.People, appears as the most important variables in most of the communities. The variable 5.year.ahead indicates the importance of forward looking, i.e. what do you expect from your community in 5 years, this is the most important driver of attachment. The variable 5.years.ago represents the perception of the community improvement, when comparing the community today with 5 years ago. Finally, the Meet.People variable indicates if the community is a good place to meet people and make friends.
\begin{figure}[hbpt]
<< import.plot3, dependson=c('data'), echo=FALSE >>=
aux <- soul.red2
aux$QSB <- with(aux, reorder(QSB, -as.numeric(QSB) ))
